---
title: "p8105_hw6_yh3824"
author: "ruby"
date: "2025-11-18"
output: github_document
---

load packages/formatting

```{r message = FALSE}
library(tidyverse)
library(mgcv)
library(modelr)
library(p8105.datasets)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Problem 1

import `homicide` data 
```{r message = FALSE} 
homicide_df = read_csv(file = "./data/homicide-data.csv")
```

Create a `city_state` variable, filtering, make resolved vs unresolved variable

```{r}
homicide_df = 
  homicide_df |> 
    mutate(
      city_state = paste(city, state, sep = ", "),
      victim_age = as.numeric(victim_age),
      resolved = if_else(str_detect(disposition, "Closed"), 1, 0) 
    ) |> 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black"),
         is.numeric(victim_age)) 
  
```

`glm` to fit logistic regression for Baltimore, MD

```{r}
baltimore_df =
  homicide_df |> 
    filter(city_state == "Baltimore, MD") 

fit = lm(resolved ~ victim_age + victim_sex + victim_race, data = baltimore_df) |> 
  broom::tidy()
  
```


write a function `lm` for all cities 

```{r}
lm_resolved = function(df) {
  
  lm(resolved ~ victim_age + victim_sex + victim_race, data = df)
}

homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  lm_resolved() |> 
  broom::tidy()
```

iterate to fit a model for each city 

```{r}
resolved_lm_results =
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    fits = map(data, lm_resolved),
    results = map(fits, ~ broom::tidy(.x, conf.int = TRUE) |> 
                    mutate(
                      odds_ratio = exp(estimate),
                      or_low = exp(conf.low),
                      or_high = exp(conf.high)
                    ))
  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, term, odds_ratio, or_low, or_high, everything())
  
```

create a plot that shows estimate ORs and CIs for each city, organize cities according to estimate OR

```{r}
resolved_lm_results |> 
  mutate(
      city_state = fct_reorder(city_state, odds_ratio), y = odds_ratio) |> 
  ggplot(aes(x = city_state, y = odds_ratio)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = or_low, ymax = or_high), width = 0.2) +
  labs(
    x = "City",
    y = "Odds Ratio",
    title = "Estimated Odds Ratio of Resolved Homicides Comparing Victim Sex"
  ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

### Comments on Plot 

xxxx


## Problem 2

import `weather_df`

```{r}
data("weather_df")
```

## Problem 3

import `birthweight` dataset 

```{r message = FALSE}
birthwgt_df = read_csv(file = "./data/birthweight.csv")
```

